---
title: "Reproducible Research. Course project1"
author: "Nadine.solov"
date: "04 12 2021"
output: html_document
---
This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data downloaded to the computer, unzipped.
 Read data into object variable dataFile

```{r,echo = TRUE }
dataFile <- read.csv("./data/activity.csv")

```

 Observe data using summary()


```{r}
summary(dataFile)
str(dataFile)

```

 Observe data again

```{r, echo=TRUE}
str(dataFile)
```
 a set of missing values using the mice package and the md.pattern function
Total number of steps taken per day
```{r}
steps_by_day <- aggregate(steps ~ date, data=dataFile, sum)
```
histogram of the total number  of steps taken each day
```{r}
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue",
     xlab="Number of Steps")
```
mean and  median number of steps  taken each day

```{r}
rmean <- mean(steps_by_day$steps)
rmean
rmedian <- median(steps_by_day$steps)
rmedian
```

Time series plot  of the average number of steps taken

```{r}

aver_numb_steps<-aggregate(steps ~ interval, data=dataFile, mean)

library(ggplot2)
ggplot(aver_numb_steps, aes(interval,steps))+
  geom_line() +
  geom_point() +
  ggtitle(expression("Average Number of Steps per Day by Interval"))+
  labs(x = "Interval", y ="Number of Steps")

```
The 5-minute interval that, on average, contains the maximum number of steps is 835:

```{r}
max_interval <- aver_numb_steps[which.max(aver_numb_steps$steps),1]
max_interval
```
 a set of missing values using the mice package and the md.pattern function

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(mice)
md.pattern(dataFile)
```
2304 values belong to a variable steps

```{r, echo=TRUE}
sum(is.na(dataFile$steps))
```
 Create a new dataset that is equal to the original dataset

```{r}
impute.dataFile <- dataFile
```
filling in missing values in the dataset
use the mean

```{r}
impute.dataFile$steps[is.na(impute.dataFile$steps)] <- mean(impute.dataFile$steps,na.rm=TRUE)
impute.dataFile$steps <- as.numeric(impute.dataFile$steps)
impute.dataFile$interval <- as.numeric(impute.dataFile$interval)
colSums(is.na(impute.dataFile))
```
Create a new dataset that is equal to the original dataset but with
the missing data filled in

```{r}
summary(impute.dataFile)
```
Make a histogram of the total number of steps taken each day

```{r}
steps_by_day_i <- aggregate(steps ~ date, impute.dataFile, sum)
hist(steps_by_day_i$steps, main = paste("Total Steps Each Day"), col="Green", 
     xlab="Number of Steps")
```
Calculate and report the mean and median total number of steps taken per day

```{r}
rmean.i <-  mean(steps_by_day_i$steps)
rmean.i
rmedian.i <- median(steps_by_day_i$steps)
rmedian.i
```

 Calculate difference between imputed and non-imputed data

```{r}
mean_diff <- rmean.i - rmean
print(mean_diff)
med_diff <- rmedian.i - rmedian
print(med_diff)
```
 Total difference

```{r}
total_diff <- sum(steps_by_day_i$steps) - sum(steps_by_day$steps)
total_diff
```
 Create a new factor variable in the dataset with two levels – “weekday” 
and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
library(chron)
table(is.weekend(impute.dataFile$date))
```
Adding new factor variable "dayofweek" indicating whether a given date is 
a weekday or weekend day

```{r}
impute.dataFile$dayofweek <- ifelse(is.weekend(impute.dataFile$date), "weekend",
                          "weekday")
table(impute.dataFile$dayofweek)
head(impute.dataFile)
```
Make a panel plot containing a time series plot (i.e. type = “l”) of
the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all weekday days or weekend days (y-axis). The plot should 

look something like the following, which was created using simulated data:
Aggregating(mean) steps over interval and day of week

```{r}
meanintervalnew<- aggregate(steps ~ interval + dayofweek, impute.dataFile, FUN=mean)
```
Aggregated Data
```{r}
head(meanintervalnew)
```
Time Series plot using ggplot

```{r}
ggplot(meanintervalnew, aes(x=interval, y=steps)) + 
  geom_line(color="red", size=1) + 
  facet_wrap(~dayofweek, nrow=2) +
  labs(x="\nInterval", y="\nNumber of steps")
```





